Summary of the things that I learned

### 2024-12-09 400-algebraic-structures.tex
Improve notation and wording

  First of all, note that for all $ n : \mathbb N $ and $ A : \Alg T $, the following diagram in $ \SET $ commutes:
  \begin{center}
    \begin{tikzcd}
      \Alg T(T_n, A) \arrow[r, "f^*"] \arrow[d, "\sim"] & \Alg{T^\prime}(f^* T_n, f^* A) \arrow[d, "\bar f_n \cdot -"]\\
      A^n \arrow[r, "\sim"] & \Alg{T^\prime}(T^\prime_n, f^* A)
    \end{tikzcd}
  \end{center}
  % \[ (g : T_n \to B) \mapsto (g: T_n \to B) \mapsto (f \cdot g : T^\prime_n \to B) \]
  % \[ (g : T_n \to B) \mapsto ((g x_i)_i : B^n) \mapsto (s \mapsto s \bullet_{f^* B} (g x_i)_i) \]
  % \[
  %   s \bullet_{f^* B} (g x_i)_i
  %   = f(s) \bullet (g x_i)_i
  %   = g(f(s) \bullet (x_i)_i)
  %   = g(f(s))
  % \]
  with $ \bar f_n: T^\prime_n \to f^* T_n $ the $ T^\prime $-algebra morphism with underlying function $ f_n $. Note that $ f^* $ is fully faithful, so the function at the top of the diagram is a bijection, and so precomposition by $ \bar f_n $ is a bijection as well. Since $ f^* $ is essentially surjective, we have for all $ A : \Alg T^\prime $, some $ B : \Alg T $ and an isomorphism $ h: f^* B \xrightarrow \sim A $, which gives the following commutative diagram:
  \begin{center}
    \begin{tikzcd}
      \Alg{T^\prime}(f^* T_n, f^* B) \arrow[d, "\bar f_n \cdot -", "\sim"'] \arrow[r, "- \cdot h"', "\sim"] & \Alg{T^\prime}(f^* T_n, A) \arrow[d, "\bar f_n \cdot -"]\\
      \Alg{T^\prime}(T^\prime_n, f^* B) \arrow[r, "- \cdot h"', "\sim"] & \Alg{T^\prime}(T^\prime_n, A)
    \end{tikzcd}
  \end{center}
  so the arrow on the right is an equivalence as well. Taking $ A = T^\prime_n $, there exists some $ g: \Alg{T^\prime}(f^* T_n, T^\prime_n) $ such that $ \bar f_n \cdot g = \id{T^\prime_n} $. Also, note that
  \[ \bar f_n \cdot g \cdot \bar f_n = \bar f_n \cdot \id{f^* T_n} \]
  so taking $ A = f^* T_n $, we see that
  \[ g \cdot \bar f_n = \id{f^* T_n}, \]
  so $ f_n $, the underlying function of $ \bar f_n $, is a bijection, with inverse $ g $.
\end{proof}

\begin{remark}
  Hyland's proof of this fact is almost the same, but in the last part, he uses some category theory. Instead of explicitly constructing the inverse, he notices that the bijection
  \[ \Alg{T^\prime}(f^* T_n, A) \xrightarrow[\bar f_n \cdot -] \sim \Alg{T^\prime}(T^\prime_n, A) \]
  is in fact the image of $ \op{\bar f_n} $ under the Yoneda embedding of $ \op{\Alg{T^\prime}} $ (or of $ \bar f_n $ under the covariant Yoneda embedding of $ \Alg{T^\prime} $) into the functor category $ \Alg{T^\prime} \to \SET $:
  \[ \yo(\bar f_n): \yo(f^* T_n) \to \yo(T^\prime_n). \]
  Since the Yoneda embedding is fully faithful, this shows that $ \bar f_n $ is an isomorphism of $ T^\prime $-algebras, so in particular, it is a bijection.
\end{remark}

@@

\begin{lemma}[\coqident{AlgebraicTheories.RepresentationTheorem}{theory_presheaf_exponentiable}]
  For all $ p $ and all $ T $-presheaves $ Q $, $ A(Q, p) $ is the exponential object $ Q^{T^p} $.
\end{lemma}
\begin{proof}
  We will show that $ A(-, p) $ constitutes a right adjoint to the functor $ - \times T^p $. We will do this using universal arrows.

@@

  Now, given a $ T $-presheaf $ Q^\prime $ and a morphism $ \psi: Q^\prime \times T^p \to Q $. Define $ \tilde \psi: Q^\prime_n \to A(Q, p)_n $ by $ \tilde \psi(q) = \psi(\iota_{n, p}(q), (x_{n + i})_i) $.

  Then $ \psi $ factors as $ (\tilde \psi \times \id{T^p}) \cdot \varphi $. Also, some equational reasoning shows that $ \tilde \psi $ is unique, which proves that $ \varphi $ indeed is a universal arrow.
\end{proof}

>>>>

  First of all, note that for all $ n : \mathbb N $ and $ A : \Alg T $, the following diagram in $ \SET $ commutes:
  \begin{center}
    \begin{tikzcd}
      \Alg T(T_n, A) \arrow[r, "f^*"] \arrow[d, "\sim"] & \Alg{T^\prime}(f^* T_n, f^* A) \arrow[d, "\overline f_n \cdot -"]\\
      A^n \arrow[r, "\sim"] & \Alg{T^\prime}(T^\prime_n, f^* A)
    \end{tikzcd}
  \end{center}
  % \[ (g : T_n \to B) \mapsto (g: T_n \to B) \mapsto (f \cdot g : T^\prime_n \to B) \]
  % \[ (g : T_n \to B) \mapsto ((g x_i)_i : B^n) \mapsto (s \mapsto s \bullet_{f^* B} (g x_i)_i) \]
  % \[
  %   s \bullet_{f^* B} (g x_i)_i
  %   = f(s) \bullet (g x_i)_i
  %   = g(f(s) \bullet (x_i)_i)
  %   = g(f(s))
  % \]
  with $ \overline f_n: T^\prime_n \to f^* T_n $ the $ T^\prime $-algebra morphism with underlying function $ f_n $. Note that $ f^* $ is fully faithful, so the function at the top of the diagram is a bijection, and so precomposition by $ \overline f_n $ is a bijection as well. Since $ f^* $ is essentially surjective, we have for all $ A : \Alg T^\prime $, some $ B : \Alg T $ and an isomorphism $ h: f^* B \xrightarrow \sim A $, which gives the following commutative diagram:
  \begin{center}
    \begin{tikzcd}
      \Alg{T^\prime}(f^* T_n, f^* B) \arrow[d, "\overline f_n \cdot -", "\sim"'] \arrow[r, "- \cdot h"', "\sim"] & \Alg{T^\prime}(f^* T_n, A) \arrow[d, "\overline f_n \cdot -"]\\
      \Alg{T^\prime}(T^\prime_n, f^* B) \arrow[r, "- \cdot h"', "\sim"] & \Alg{T^\prime}(T^\prime_n, A)
    \end{tikzcd}
  \end{center}
  so the arrow on the right is an equivalence as well. Taking $ A = T^\prime_n $, there exists some $ g: \Alg{T^\prime}(f^* T_n, T^\prime_n) $ such that $ \overline f_n \cdot g = \id{T^\prime_n} $. Also, note that
  \[ \overline f_n \cdot g \cdot \overline f_n = \overline f_n \cdot \id{f^* T_n} \]
  so taking $ A = f^* T_n $, we see that
  \[ g \cdot \overline f_n = \id{f^* T_n}, \]
  so $ f_n $, the underlying function of $ \overline f_n $, is a bijection, with inverse $ g $.
\end{proof}

\begin{remark}
  Hyland's proof of this fact is almost the same, but in the last part, he uses some category theory. Instead of explicitly constructing the inverse, he notices that the bijection
  \[ \Alg{T^\prime}(f^* T_n, A) \xrightarrow[\overline f_n \cdot -] \sim \Alg{T^\prime}(T^\prime_n, A) \]
  is in fact the image of $ \op{\overline f_n} $ under the Yoneda embedding of $ \op{\Alg{T^\prime}} $ (or of $ \overline f_n $ under the covariant Yoneda embedding of $ \Alg{T^\prime} $) into the functor category $ \Alg{T^\prime} \to \SET $:
  \[ \yo(\overline f_n): \yo(f^* T_n) \to \yo(T^\prime_n). \]
  Since the Yoneda embedding is fully faithful, this shows that $ \overline f_n $ is an isomorphism of $ T^\prime $-algebras, so in particular, it is a bijection.
\end{remark}

@@

\begin{lemma}[\coqident{AlgebraicTheories.RepresentationTheorem}{theory_presheaf_exponentiable}]\label{lem:plus-p-presheaf}
  For all $ p $ and all $ T $-presheaves $ Q $, $ A(Q, p) $ is the exponential object $ Q^{T^p} $.
\end{lemma}
\begin{proof}
  We will show that $ A(-, p) $ constitutes a right adjoint to the functor $ - \times T^p $. We will do this using universal arrows.

@@

  Now, given a $ T $-presheaf $ Q^\prime $ and a morphism $ \psi: Q^\prime \times T^p \to Q $. Define $ \widetilde \psi: Q^\prime_n \to A(Q, p)_n $ by $ \widetilde \psi(q) = \psi(\iota_{n, p}(q), (x_{n + i})_i) $.

  Then $ \psi $ factors as $ (\widetilde \psi \times \id{T^p}) \cdot \varphi $. Also, some equational reasoning shows that $ \widetilde \psi $ is unique, which proves that $ \varphi $ indeed is a universal arrow.
\end{proof}

### 2024-12-09 500-previous-work.tex
Improve notation and wording

  To show that $ (\alpha \cdot -) \dashv \alpha_* $, we need to show that for all $ Y : \R^B $, $ (\alpha_* Y, \epsilon_Y) $ is a universal arrow from $ (\alpha \cdot -) $ to $ Y $. Recall the following diagram:
  \begin{center}
    \begin{tikzcd}
      Z \circ \alpha \arrow[rd, "f"'] \arrow[r, "{\hat f \circ \alpha}", dashed] & (\alpha_* Y) \circ \alpha \arrow[d, "{\epsilon_Y}"]\\
      & Y
    \end{tikzcd}
  \end{center}
  Now, suppose that we have some $ Z : \R^A $ and some $ f: \R^B(Z \circ \alpha, Y) $. We define
  \[ \hat f = \lambda x_1 x_2 x_3, f(g^{-1}(x_1, x_3)) (Z x_1 x_2): \R^A(Z, \alpha_* Y) \]
  and we have $ ((\hat f) \circ \alpha) \cdot \epsilon_Y = f $.
  % \begin{align*}
  %   (((\alpha \cdot -) \hat f) \cdot \epsilon_Y) b z
  %   &= \epsilon_Y b ((\hat f (\alpha b)) z)\\
  %   &= \epsilon_Y b (\lambda x_3, f (g^{-1} ((\alpha b), x_3)) (Z (\alpha b) z))\\
  %   &= Y b ((\lambda x_3, f (g^{-1} ((\alpha b), x_3)) (Z (\alpha b) z)) (\pi_2 (g b)))\\
  %   &= Y b (f (g^{-1} ((\alpha b), \pi_2 (g b))) (Z (\alpha b) z))\\
  %   &= Y b (f (g^{-1} (g b)) (Z (\alpha b) z))\\
  %   &= f b z.
  % \end{align*}
  Now, suppose that we have some (other) $ \hat f^\prime : \R^A(Z, \alpha_*(Y)) $ such that $ (\hat f^\prime \circ \alpha) \cdot \epsilon_Y = f $. Then substituting $ (\hat f^\prime \circ \alpha) \cdot \epsilon_Y $ for $ f $ in $ \hat f $ yields
  \[ \hat f = \lambda x_1, (\alpha_* Y x_1) \circ (\hat f^\prime x_1) = \hat f^\prime, \]
  so $ \hat f $ is unique and $ (\alpha_* Y, \epsilon_Y) $ is a universal arrow, which concludes the proof that $ (\alpha \cdot -) \dashv \alpha_* $ and we see that $ \R $ is cartesian closed relative to the given class of display maps.
  % \begin{align*}
  %   \hat f
  %   &= \lambda x_1 x_2 x_3, f (g^{-1} (x_1, x_3))(Z x_1 x_2)\\
  %   &= \lambda x_1 x_2 x_3, ((\epsilon_Y (g^{-1} (x_1, x_3))) \circ ((\hat f^\prime \circ \alpha) (g^{-1} (x_1, x_3)))) (Z x_1 x_2)\\
  %   &= \lambda x_1 x_2 x_3, (\epsilon_Y (g^{-1} (x_1, x_3))) (\hat f^\prime (\alpha (g^{-1} (x_1, x_3))) (Z x_1 x_2))\\
  %   &= \lambda x_1 x_2 x_3, (\epsilon_Y (g^{-1} (x_1, x_3))) (\hat f^\prime x_1 x_2)\\
  %   &= \lambda x_1 x_2 x_3, Y (g^{-1} (x_1, x_3)) (\hat f^\prime x_1 x_2 (\pi_2 (g (g^{-1} (x_1, x_3)))))\\
  %   &= \lambda x_1 x_2 x_3, Y (g^{-1} (x_1, x_3)) (\hat f^\prime x_1 x_2 (X x_1 x_3))\\
  %   &= \lambda x_1 x_2, \alpha_* Y x_1 (\hat f^\prime x_1 (Z x_1 x_2))\\
  %   &= \hat f^\prime
  % \end{align*}
\end{proof}

@@

Now, it is not hard to show that Hyland's and Selinger's notions of $ \lambda $-theory are equivalent: If we have one of Selinger's $ \lambda $-theories with a countable set of variables, enumerate the variables as $ x_1, x_2, \dots $ and let $ L_n $ be the set of terms with only the free variables $ x_1, \dots, x_n $. Then, if we quotient by the equivalence relation, we get one of Hyland's $ \lambda $-theories. Conversely, if we start with one of Hyland's $ \lambda $-theories $ L $, we can consider $ L_n $ as a subset of $ L_{n + 1} $ by sending $ x_{n, i} $ to $ x_{n + 1, i} $. If we take the union $ \bar L = \bigcup_n L_n $ (officially, this is a directed limit), we have a function $ f: \Lambda_{\bar L} \to \bar L $, because the constants of $ \Lambda_{\bar L} $ are exactly the terms of $ \bar L $ and because $ \lambda $-theories in Hyland's sense support the operations of the $ \lambda $-calculus (see Subsection \ref{subsec:lambda-calculus-operations}). This gives an equivalence relation on $ \Lambda_{\bar L} $ by taking $ a\ `='\ b $ if $ f(a) = f(b) $.

>>>>

  To show that $ (\alpha \cdot -) \dashv \alpha_* $, we need to show that for all $ Y : \R^B $, $ (\alpha_* Y, \epsilon_Y) $ is a universal arrow from $ (\alpha \cdot -) $ to $ Y $. Recall the following diagram:
  \begin{center}
    \begin{tikzcd}
      Z \circ \alpha \arrow[rd, "f"'] \arrow[r, "{\widehat f \circ \alpha}", dashed] & (\alpha_* Y) \circ \alpha \arrow[d, "{\epsilon_Y}"]\\
      & Y
    \end{tikzcd}
  \end{center}
  Now, suppose that we have some $ Z : \R^A $ and some $ f: \R^B(Z \circ \alpha, Y) $. We define
  \[ \widehat f = \lambda x_1 x_2 x_3, f(g^{-1}(x_1, x_3)) (Z x_1 x_2): \R^A(Z, \alpha_* Y) \]
  and we have $ ((\widehat f) \circ \alpha) \cdot \epsilon_Y = f $.
  % \begin{align*}
  %   (((\alpha \cdot -) \widehat f) \cdot \epsilon_Y) b z
  %   &= \epsilon_Y b ((\widehat f (\alpha b)) z)\\
  %   &= \epsilon_Y b (\lambda x_3, f (g^{-1} ((\alpha b), x_3)) (Z (\alpha b) z))\\
  %   &= Y b ((\lambda x_3, f (g^{-1} ((\alpha b), x_3)) (Z (\alpha b) z)) (\pi_2 (g b)))\\
  %   &= Y b (f (g^{-1} ((\alpha b), \pi_2 (g b))) (Z (\alpha b) z))\\
  %   &= Y b (f (g^{-1} (g b)) (Z (\alpha b) z))\\
  %   &= f b z.
  % \end{align*}
  Now, suppose that we have some (other) $ \widehat f^\prime : \R^A(Z, \alpha_*(Y)) $ such that $ (\widehat f^\prime \circ \alpha) \cdot \epsilon_Y = f $. Then substituting $ (\widehat f^\prime \circ \alpha) \cdot \epsilon_Y $ for $ f $ in $ \widehat f $ yields
  \[ \widehat f = \lambda x_1, (\alpha_* Y x_1) \circ (\widehat f^\prime x_1) = \widehat f^\prime, \]
  so $ \widehat f $ is unique and $ (\alpha_* Y, \epsilon_Y) $ is a universal arrow, which concludes the proof that $ (\alpha \cdot -) \dashv \alpha_* $ and we see that $ \R $ is cartesian closed relative to the given class of display maps.
  % \begin{align*}
  %   \widehat f
  %   &= \lambda x_1 x_2 x_3, f (g^{-1} (x_1, x_3))(Z x_1 x_2)\\
  %   &= \lambda x_1 x_2 x_3, ((\epsilon_Y (g^{-1} (x_1, x_3))) \circ ((\widehat f^\prime \circ \alpha) (g^{-1} (x_1, x_3)))) (Z x_1 x_2)\\
  %   &= \lambda x_1 x_2 x_3, (\epsilon_Y (g^{-1} (x_1, x_3))) (\widehat f^\prime (\alpha (g^{-1} (x_1, x_3))) (Z x_1 x_2))\\
  %   &= \lambda x_1 x_2 x_3, (\epsilon_Y (g^{-1} (x_1, x_3))) (\widehat f^\prime x_1 x_2)\\
  %   &= \lambda x_1 x_2 x_3, Y (g^{-1} (x_1, x_3)) (\widehat f^\prime x_1 x_2 (\pi_2 (g (g^{-1} (x_1, x_3)))))\\
  %   &= \lambda x_1 x_2 x_3, Y (g^{-1} (x_1, x_3)) (\widehat f^\prime x_1 x_2 (X x_1 x_3))\\
  %   &= \lambda x_1 x_2, \alpha_* Y x_1 (\widehat f^\prime x_1 (Z x_1 x_2))\\
  %   &= \widehat f^\prime
  % \end{align*}
\end{proof}

@@

Now, it is not hard to show that Hyland's and Selinger's notions of $ \lambda $-theory are equivalent: If we have one of Selinger's $ \lambda $-theories with a countable set of variables, enumerate the variables as $ x_1, x_2, \dots $ and let $ L_n $ be the set of terms with only the free variables $ x_1, \dots, x_n $. Then, if we quotient by the equivalence relation, we get one of Hyland's $ \lambda $-theories. Conversely, if we start with one of Hyland's $ \lambda $-theories $ L $, we can consider $ L_n $ as a subset of $ L_{n + 1} $ by sending $ x_{n, i} $ to $ x_{n + 1, i} $. If we take the union $ \overline L = \bigcup_n L_n $ (officially, this is a directed limit), we have a function $ f: \Lambda_{\overline L} \to \overline L $, because the constants of $ \Lambda_{\overline L} $ are exactly the terms of $ \overline L $ and because $ \lambda $-theories in Hyland's sense support the operations of the $ \lambda $-calculus (see Subsection \ref{subsec:lambda-calculus-operations}). This gives an equivalence relation on $ \Lambda_{\overline L} $ by taking $ a\ `='\ b $ if $ f(a) = f(b) $.
